# (PART\*) PhD Program {.unnumbered}

# PhD Milestones {#phdprogram}

There are four major milestones in the PhD Program, as diagrammed below. [Coursework](#phdcoursework) leads directly to the [Qualfiying Papers](#qp) requirement, which must be completed before the [Prospectus](#prospectus) and [Dissertation].

```{r, echo=FALSE}
DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = LR]
  
  node [shape = rectangle]        
  rec1 [label = 'Coursework']
  rec2 [label = 'Qualifying Papers']
  rec3 [label =  'Prospectus']
  rec4 [label = 'Dissertation']
  
  # edge definitions with the node IDs
  rec1 -> rec2 -> rec3 -> rec4
  }",
  height = 100)
```

In order to make adequate progress toward degree and maintain good standing in the program, PhD students are expected to meet the following degree milestones.

| Milestone         | Deadline            |
|:------------------|:--------------------|
| Complete QP1      | end of 4th semester |
| Complete QP2      | end of 6th semester |
| Defend prospectus | end of 7th semester |

Failure to meet these milestones by the stipulated deadline may result in departmental warning of [academic probation](#probation), jeopardizing eligibility for departmental funding and risking dismissal from the program.

The following is an outline of typical activities for PhD by year in the program. Individual timelines may vary based on a number of factors, including previous training in linguistics and previous research and community experience.

**Year 1**

-   foundational coursework
-   meet with faculty members to discuss ideas for QP1 and QP2
-   Departmental participation (reading group, LSM, LDTC, etc.)
-   prepare proposal for summer research

The first year of the PhD program will be the most coursework-intensive year of your studies. Depending on what courses you have been exempted from, you will likely be enrolled in foundational courses in the core area of linguistics. However, it is important to bear in mind that the PhD is a research-oriented degree. While taking courses during the first year, you should be thinking about idea for your first [Qualifying Paper](#qp). QP's typically arise out of a class paper, and you may want to discuss with your instructor whether a particular paper might serve as the basis for a QP. This is also a good time to ask whether that instructor would be willing to supervise the QP.

First-year students are also encouraged to get involved with one or more of the many student-led groups, such as Reading Groups and the Linguistic Society of MÄnoa. This is not only a great way to get to know your colleagues, it's a great way to be a good citizen of the department.

The summer after your first year is a great time to begin exploratory field work or begin data collection for a QP. Applications for summer research funding are circulated in the Spring. Note that QP1 must be complete by the end of your 4th semester, so this first summer will be your only summer break available for data collection for QP1.

**Year 2**

-   complete required coursework
-   possible directed research (LING 699)
-   propose and complete QP1
-   propose QP2
-   departmental service (reading group leader, LSM officer, etc.)

During the second year you will likely need to take some additional foundational courses, but all required courses should be complete by the end of the year. Once all coursework requirements are complete you may enroll in [Directed Research (LING 699)](#ling699) in order to focus on your QP1. Plan to hold the consultation meeting for QP1 in the Fall and to complete your QP1 in the Spring.

This is also a good time to begin thinking about your second Qualifying Paper and identifying a faculty member who will supervise that paper.

The second year is a good time to take more of a leadership role in service activities, perhaps by leading a Reading Group or serving as an LSM officer.

**Year 3**

-   possible elective coursework
-   directed reading (LING 699)
-   complete QP2
-   begin drafting dissertation prospectus
-   exploratory data collection (field work, experimentation)
-   form PhD advisory committee
-   possible funding applications for data collection

The third year represents a transition from coursework to research. If you do choose to enroll in elective courses, you should choose to audit or take the course credit/no-credit. This often presents students with a very difficult choice, namely, the choice to forego an interesting elective course in order to focus on research. It is critical that these electives do not interfere with or detract from your research this year. Consider enrolling in an elective course *only* if it is directly relevant to your QP2 or your dissertation.

Your main objectives for this year are to (i) complete QP2; and (ii) begin drafting your [dissertation prospectus](#Dissertation%20Proposal%20and%20Prospectus). You will likely enroll in Directed Research courses (LING 699) to facilitate this work.

As your research workload increases, this is a good time to start scaling back on service activities. It's okay (and encouraged) to keep participating in student organizations, but this is a good time to let newer cohorts take over leadership roles so that you can focus on your research.

If you anticipate needing additional funding to support your dissertation research, this is a good time to prepare applications to external funders such as NSF and ELDP. Those applications very much parallel the prospectus, so can often be drafted in tandem.

By the end of the third year you should have a good sense of faculty research interests and have identified a faculty member who is willing to supervise your dissertation research. This may or may not be the person initially assigned to advise you. Once you have identified a dissertation supervisor, work with them to identify other appropriate members for your PhD advisory committee.

The summer after your third year is often a critical time for data collection and field work. You will be in the process of drafting your dissertation prospectus, and the summer can be used to consult with community members about your proposed work and to gather additional data to demonstrate the feasibility of your dissertation plans.

**Year 4**

-   defend prospectus and advance to candidacy
-   LING 800
-   data collection and analysis
-   begin drafting dissertation

This is the year in which you will advance to candidacy and become a PhD "candidate," also known as ABD (all but dissertation). You no longer need to enroll in courses but instead register for 1 credit of LING 800 (dissertation research). You are now a full-time researcher. In particular, you should NOT be taking courses, either in Linguistics or in other departments. Your dissertation now becomes your primary academic responsibility.

You may want to start drafting some chapters of your dissertation while you are engaged in data collection and analysis.

You may also find that you need to spend some time revising QP's which have been submitted for publication. Be careful not to let this activity detract from your dissertation research.

**Year 5**

-   continued data collection, if relevant
-   LING 800
-   dissertation writing
-   dissertation defense
-   publish dissertation research in journals
-   prepare job applications

During the fifth year you will shift your focus from research to writing. As with the transition from coursework to research, this is often a difficult transition for students. Particularly with language documentation and other types of community-based research, it can be difficult to know when the research is "finished." Questions will arise during the dissertation writing process which beg for additional data or will tempt you to seek input for community collaborators. Yet at some point you will need to draw a line and refrain from returning to data collection mode, opting to write up the analysis based on the data at hand. Knowing where and when to draw this line can be difficult, and you are encouraged to discuss these issues early and often with your advisor.

The earlier you defend your dissertation, the more time you will have for preparing journal articles derived from the dissertation and preparing job applications. Don't wait until May of your fifth year to defend!

**Beyond Year 5**

We recognize that circumstances do not always allow students to complete degree requirements within this 5-year timeline. We are committed to supporting PhD students through to completion of their degrees, and additional funding *may* sometimes be available. However, to be eligible for funding beyond the 5th year, students must demonstrate continued progress toward degree and provide evidence of extraordinary circumstances.

# PhD Coursework {#phdcoursework}

Coursework is the first stage of the PhD Program. All students in the PhD program are required to complete a minimum of 33 credit hours of course work at the University of Hawai'i *beyond courses counted towards the MA degree*. This course work must include LING 621 (Phonology), LING 622 (Grammar), and two Methods courses. For students transferring from the MA program, the MA "core" courses (LING 410, 420, 421, 422 and possibly 645, depending on MA stream) cannot be applied toward the PhD. For students who have already earned an MA in Linguistics from UH, none of the 30 credits applied toward the MA degree may also be counted toward the PhD.

A student may be exempted from a 600-level course such as LING 621 and 622 with instructor approval. Such exemptions are granted only when it is clear that the student has taken a closely-overlapping course at another institution, and that the student still maintains control over the relevant material. In such cases, the student's 33 credit requirement is reduced by 3 credits (per exempted course).

Students interested in experimental research are strongly advised to take one or more courses in statistical analysis (e.g. LING 640G: Statistics in Linguistics, EDEP 429, SLS 490, SLS 671).

LING 750G Professional Development (ICLDC Prep Course) may be taken multiple times, but will only be counted once towards a degree:

i.  the course will count for a maximum of 3 credits towards the PhD 33 credit requirement, and
ii. if it is used to satisfy an MA degree requirement, subsequent LING 750G courses may not be used towards the PhD degree.

Students who have not already taken or received a course exemption for the following 'core' courses must take these courses *in addition* to the 33 credit hours of course work required of PhD students.

-   LING 410 Articulatory Phonetics
-   LING 420 Morphology
-   LING 421 Introduction to Phonological Analysis
-   LING 422 Introduction to Grammatical Analysis
-   LING 645 The Comparative Method

If you already have advanced coursework in linguistics, you may elect to enroll in a [Directed Research](#ling699) course.

At the discretion of the department faculty or your advisor, you may be required to take [additional writing courses](#writing). A description of adequate writing skills and a list of courses and their descriptions can be found under the Adequate Writing Skills section of this document. This information is also available to download from our department website at

## Breadth Requirement {#breadth}

In order to ensure that our graduates have sufficient breadth of knowledge, the department requires that each student have multiple areas of specialization. Within the 33 credit hours of course work, students must take at least three graduate-level courses in each of two distinct areas of linguistics. In addition, students are strongly encouraged to gain expertise in a third area of linguistics for the purpose of employment and further research opportunities.

```{r, echo=FALSE}
DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = TB]
  
  node [shape = oval,fontsize=10]
  area [label = 'Specialization Area']

  node [shape=rect,fixedsize=true,width=1.5,fontsize=10]
  course1 [label='Course 1\nletter grade A-']
  course2 [label='Course 2\nletter grade A-']
  course3 [label='Course 3\ngrade A- or CR\n(including LING 699']

  # edge definitions with the node IDs
  area -> course1
  area -> course2
  area -> course3
  }",
  height=200)
```

Various areas of linguistics are suitable for a specializaiton, including a specialization in a particular language or language area. Two-semester courses sequences will count as two courses. Graduate-level courses taken as an undergraduate or as an MA student and courses taken at other universities may count as part of a speicalization area, with approval from both the student's advisor and the Graduate Chair. Some 400-level courses may count as part of a specialization area, with advisor approval. A maximum of three credits per area may be taken as 699, if the student requests and receives approval by the Graduate Chair in writing. At least two of the three classes in each focus area must be taken for a letter grade, with a minimum grade requirement of A-. If your third class is taken as credit/no credit, you must receive a grade of CR; if your third class is taken for a letter grade, you must receive a grade of C or better; you cannot audit this third course.

The specialization areas must be approved by your advisor. The advisor (in consultation with the Graduate Chair if necessary), will work to ensure that each area represents a coherent set of courses, and that the areas are sufficiently different and non-overlapping to satisfy the breadth requirement. You will need to consult early with your advisor regarding the suitability of the areas of courses you plan on.

There will be courses that may be counted in one area or another, but no course can be counted more than once. Students should consult the instructor, their advisor, and the Graduate Chair about the area assignment of any course whose status is not obvious.

## Course Exemption

If you wish to seek exemption from any of the four core MA courses, you must meet with the faculty member in charge of exemption for the course on the Monday or Tuesday of the week prior to the first day of instruction. During this meeting you must provide the instructor with the syllabus from your prior course, and be prepared to discuss course content and possibly be quizzed on course materials. The instructor will then decide to

-   exempt you from the class, or
-   not exempt you from the class, or
-   require you to take an exemption exam for the course (in this case a minimum passing grade is B (not B-)), or
-   require you to audit the entire class or a portion of the class; the instructor will inform you of requirements for a successful audit.

You must notify the Department secretary no later than August 1st (for Fall) and December 1st (for Spring). The secretary will coordinate a meeting with the faculty member in charge of the exemption.

Exemption exams, if deemed required by the instructor, are held during the week prior to the first day of instruction and are scheduled in two hour blocks. If you are required to take one or more of these exams, you must notify the Department secretary immediately after your meeting with the faculty member in charge of the exemption which exam(s) you are going to take. The secretary will schedule your exemption exam(s). All exemption exams must be completed no later than Wednesday of the first week of classes. If you are required to take the exemption exam you should attend the course until your exam has been graded and an exemption has been granted.

Instructors must provide exemption documentation or exam grades to the Graduate Chair and relevant students no later than the day before the last day to add classes. You may take any given exemption exam only once. Should you not pass an exam, you must take the relevant course at the first availability, i.e., the same semester if the course is offered (if not, the following semester).

If you intend to seek exemption(s), you must make every effort to do so within the first year of your program.

## Directed Research (LING 699) {#ling699}

Once you have completed the required PhD coursework, you may register for a Directed Research course (LING 699). Directed Research allows a student to pursue research for a Qualifying Paper, Dissertation Proposal or other project---under the direction of an appropriate faculty member. Students wishing to register for LING 699 should discuss the possibility with the proposed instructor, recognizing that faculty members may not always be available for supervising these courses. If the faculty member agrees to supervise your directed research, complete a [LING 699 Proposal](files/699form.pdf), including the following information:

1.  the scope of the proposed course, with references
2.  the relation of the proposed research to your degree program and career goals
3.  timetable of various components within the semester
4.  a brief description of the final product(s) of the course (e.g., QP draft, proposal draft, etc.)
5.  number of credits (maximum 6)

Have your instructor sign the form and then submit the form to the department office. You will then be assigned a course registration number so that you can register for the course. It is best to discuss the proposed LING 699 during the registration period at the end of the previous semester, but the form can be completed during the first week of the term.

## Advising Record and Annual Student Evaluation

One tool to help you to track your progress towards your degree is your advising record. A copy of your latest record is placed in your student mailbox at the beginning of each Fall semester. Review it and inform the department office of any discrepancies so that they can be updated.

All-but-Dissertation (ABD) students will not receive a copy of their advising record because these students have already completed all requirements except for your dissertation. A sample PhD Advising Record can be found in the [Forms] section of the handbook.

At the end of each academic year, the Linguistics faculty holds its annual student evaluation meeting during which the progress of every student in the department is discussed. Your advisor will be provided with copies of your current advising record and a semester report of your grades. During the meeting your advising record is updated based on the completion of courses and degree requirements during the past academic year. Note is also made of exceptionally fast or slow progress.

A student with unsatisfactory progress will be contacted by his/her advisor and/or the Grad Chair and may be put on warning of academic probation, or actually placed on probation. (See probation section for more information). The student will be required to acknowledge receipt of notification of this unsatisfactory progress by signing the Graduate Student Annual Progress Form.
